---
title: "writing_functions"
output: github_document
    
---


```{r setup, include=FALSE}
library(tidyverse)
library(p8105.datasets)
library(rvest)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## My first function

```{r}
x_vec = rnorm(25, mean = 5, sd = 3)

(x_vec - mean(x_vec)) / sd(x_vec)

```

```{r}
z_scores = function(x) {
  z = (x-mean(x))/sd(x)
  z
}
z_scores(x_vec)
```

Try my function on some other things.These should give errors. 

```{r, error=TRUE}
z_scores(3)
z_scores("my name is eman")
z_scores(mt_cars)
z_scores(c(TRUE, TRUE, FALSE, TRUE))
```

Making your function safe

```{r}
z_scores = function(x) {
  
  if(!is.numeric(x)){
    stop ("Input must be numeric")
  }
  
  if(length(x) ==1){
    stop("Z scores cannot be computed for length 1 vectors")
  }
  
  z = (x-mean(x)) / sd(x)
  z
}
```



## Multiple Outputs

Return a list

```{r}
mean_and_sd = function(x){
  if(!is.numeric(x)) stop ("Argument x should be numeric")
  if(length(x) ==1) stop ("Cannot be computed for length 1 vectors")
  
  list(mean = mean(x), sd = sd(x))
}
mean_and_sd(x_vec)
```

Return a tibble (a 1-row table)

```{r}
mean_and_sd = function(x){
  if(!is.numeric(x)) stop ("Argument x should be numeric")
  if(length(x) ==1) stop ("Cannot be computed for length 1 vectors")
  
  tibble(mean=mean(x), sd = sd(x))
}
mean_and_sd(x_vec)
```

## Multiple inputs + simulation
Goal: Make a function that creates data and summarizes it.

Try it once (no function)
```{r}
sim_data = tibble(x=rnorm(30, mean = 2, sd = 3))

sim_data |> 
  summarize(mu_hat = mean(x), sigma_hat = sd(x))

```
Wrap into a function argument 
```{r}
sim_mean_sd = function(n, mu=2, sigma = 3){
  sim_data = tibble(x=rnorm(n, mean = mu, sd = sigma))

sim_data |> 
  summarize(mu_hat = mean(x), sigma_hat = sd(x))
}
sim_mean_sd(30)
```

LoTR Excel example. The "old" way
```{r}
fellowship_ring = readxl::read_excel("LotR_Words.xlsx", range="B3:D6") |> mutate(movie="fellowship_ring")
two_towers      = readxl::read_excel("LotR_Words.xlsx", range="F3:H6") |> mutate(movie="two_towers")
return_king     = readxl::read_excel("LotR_Words.xlsx", range="J3:L6") |> mutate(movie="return_king")

lotr_tidy = bind_rows(fellowship_ring, two_towers, return_king) |>
  janitor::clean_names() |>
  pivot_longer(female:male, names_to="sex", 
values_to="words") |>
  mutate(race = str_to_lower(race)) |>
  select(movie, everything())

```

## Learning Assessment
The "better" way.

```{r}
lotr_load_and_tidy = function(path, range, movie_name) {
  readxl::read_excel(path, range = range) |>
    janitor::clean_names() |>
    pivot_longer(female:male, names_to = "sex", values_to = "words") |>
    mutate(race = str_to_lower(race), movie = movie_name) |>
    select(movie, everything())
}

lotr_tidy =
  bind_rows(
    lotr_load_and_tidy("LotR_Words.xlsx", "B3:D6", "fellowship_ring"),
    lotr_load_and_tidy("LotR_Words.xlsx", "F3:H6", "two_towers"),
    lotr_load_and_tidy("LotR_Words.xlsx", "J3:L6", "return_king")
  )

```


## Web scraping example (NSSDUH data)

1. Get the webpage

```{r}
nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

nsduh_html = read_html(nsduh_url)
```

2. Clean one table (no function yet)

```{r}
data_marj =
  nsduh_html |>          
  html_table() |>        
  nth(1) |>              
  slice(-1) |>           
  select(-contains("P Value")) |>  
  pivot_longer(-State, names_to="age_year", values_to="percent") |>
  separate(age_year, into=c("age","year"), sep="\\(") |>
  mutate(
    year    = str_replace(year, "\\)", ""),
    percent = str_replace(percent, "[a-c]$", ""), 
    percent = as.numeric(percent)
  ) |>
  filter(!(State %in% c("Total U.S.","Northeast","Midwest","South","West")))

```

3. Wrap it into a function so you can reuse it

```{r}
nsduh_table = function(html, table_num, table_name) {
  html |>
    html_table() |>
    nth(table_num) |>
    slice(-1) |>
    select(-contains("P Value")) |>
    pivot_longer(-State, names_to="age_year", values_to="percent") |>
    separate(age_year, into=c("age","year"), sep="\\(") |>
    mutate(
      year    = str_replace(year, "\\)", ""),
      percent = str_replace(percent, "[a-c]$", ""),
      percent = as.numeric(percent),
      name    = table_name
    ) |>
    filter(!(State %in% c("Total U.S.","Northeast","Midwest","South","West")))
}

```

4. Use it for multiple tables easily

```{r}
nsduh_results =
  bind_rows(
    nsduh_table(nsduh_html, 1, "marj_one_year"),
    nsduh_table(nsduh_html, 4, "cocaine_one_year"),
    nsduh_table(nsduh_html, 5, "heroin_one_year")
  )

```

## Passing a function into a function

```{r}
my_summary = function(x, summ_func){
  summ_func(x)
}
x_vec=rnorm(25, 0, 1)
my_summary(x_vec, mean)
my_summary(x_vec, sd)
my_summary(x_vec, IQR)
my_summary(x_vec, var)

```
## Scoping (how R finds variable names)

```{r}
f = function(x){
  z = x+y
  z
}
x=1
y=2
f(x=y)
```
Let's define x

```{r}
f = function(x){
  z = x+y
  z
}
x=1
y=2
f(x)
```














